HBCC - DLPFC - FEATURECOUNTS - GRCh38 - CQN

### 2020-05-14 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
add hbcc dlpfc cohort

knit2synapse::createAndKnitToFolderEntity(file = "HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                          parentId ="syn21435176",
                                          folderName = 'HBCC - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
```

@@

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(mvBIC)

@@

library(variancePartition)
library(synapser)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn21435176';
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (DLPFC)';

@@

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath = thisFileName)
```
### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

@@

# Stepwise regression
phase1 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = base_formula,
                         data = COVARIATES,
                         variables = array(c("(1|Reported_Gender)", "scale(PMI)",
                                             "scale(ageOfDeath)", "scale(RIN)", "scale(RIN2)", "scale(EV.1)",
                                             "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))
phase1
```

#### Phase II (batch effects)
```{r phase2}
# Fit a base model with the following covariates
phase2 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase1$formula, 
                         data = COVARIATES,
                         variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
phase2
```

#### Phase III (RNASeq alignment specific covariates)
```{r phase3}
# Fit a base model with the following covariates
phase3 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase2$formula, 
                         data = COVARIATES,
                         variables = array(c("scale(IntragenicRate)", "scale(IntronicRate)","scale(IntergenicRate)","scale(rRNARate)", 
                                             "scale(TotalReads)", "scale(GenesDetected)", "scale(MappedReads)")))
phase3
```

@@

The model for clinical, technical and batch variables selected after phase I, II and III is `r paste(phase3$formula)`

>>>>

knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                          parentId ="syn22012942",
                                          folderName = 'HBCC - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
```

@@

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

@@

library(variancePartition)
library(synapser)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
library(mvIC) # devtools::install_github("GabrielHoffman/mvIC", repos=BiocManager::repositories())

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22012942';
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (DLPFC)';

@@

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
```
### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

@@

# Stepwise regression
phase1 <- mvIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = base_formula,
                         data = COVARIATES,
                         variables = array(c("(1|Reported_Gender)", "scale(PMI)",
                                             "scale(ageOfDeath)", "scale(RIN)", "scale(RIN2)", "scale(EV.1)",
                                             "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))
kable(phase1$settings)
kable(phase1$trace)
as.character(phase1$formula)[2]
```

#### Phase II (batch effects)
```{r phase2}
# Fit a base model with the following covariates
phase2 <- mvIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase1$formula, 
                         data = COVARIATES,
                         variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
kable(phase2$settings)
kable(phase2$trace)
as.character(phase2$formula)[2]
```

#### Phase III (RNASeq alignment specific covariates)
```{r phase3}
# Fit a base model with the following covariates
phase3 <- mvIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase2$formula, 
                         data = COVARIATES,
                         variables = array(c("scale(IntragenicRate)", "scale(IntronicRate)","scale(IntergenicRate)","scale(rRNARate)", 
                                             "scale(TotalReads)", "scale(GenesDetected)", "scale(MappedReads)")))
kable(phase3$settings)
kable(phase3$trace)
as.character(phase3$formula)[2]
```

@@

The model for clinical, technical and batch variables selected after phase I, II and III is:
```{r selection_results}
as.character(phase3$formula)[2]
```
